---
title: "Class15"
author: "Nicholas Thiphakhinkeo A17686679"
date: "11/19/24"
format: pdf
---
> Q1. datapasta to assign CDC pertussis to df + ggplot (cases/time)

# [CDC data](https://www.cdc.gov/pertussis/php/surveillance/pertussis-cases-by-year.html?CDC_AAref_Val=https://www.cdc.gov/pertussis/surv-reporting/cases-by-year.html)
```{r}
cdc <- data.frame(
                                 year = c(1922L,1923L,1924L,1925L,
                                          1926L,1927L,1928L,1929L,1930L,1931L,
                                          1932L,1933L,1934L,1935L,1936L,
                                          1937L,1938L,1939L,1940L,1941L,1942L,
                                          1943L,1944L,1945L,1946L,1947L,
                                          1948L,1949L,1950L,1951L,1952L,
                                          1953L,1954L,1955L,1956L,1957L,1958L,
                                          1959L,1960L,1961L,1962L,1963L,
                                          1964L,1965L,1966L,1967L,1968L,1969L,
                                          1970L,1971L,1972L,1973L,1974L,
                                          1975L,1976L,1977L,1978L,1979L,1980L,
                                          1981L,1982L,1983L,1984L,1985L,
                                          1986L,1987L,1988L,1989L,1990L,
                                          1991L,1992L,1993L,1994L,1995L,1996L,
                                          1997L,1998L,1999L,2000L,2001L,
                                          2002L,2003L,2004L,2005L,2006L,2007L,
                                          2008L,2009L,2010L,2011L,2012L,
                                          2013L,2014L,2015L,2016L,2017L,2018L,
                                          2019L,2020L,2021L,2022L,2024L),
         cases = c(107473,164191,165418,152003,
                                          202210,181411,161799,197371,
                                          166914,172559,215343,179135,265269,
                                          180518,147237,214652,227319,103188,
                                          183866,222202,191383,191890,109873,
                                          133792,109860,156517,74715,69479,
                                          120718,68687,45030,37129,60886,
                                          62786,31732,28295,32148,40005,
                                          14809,11468,17749,17135,13005,6799,
                                          7717,9718,4810,3285,4249,3036,
                                          3287,1759,2402,1738,1010,2177,2063,
                                          1623,1730,1248,1895,2463,2276,
                                          3589,4195,2823,3450,4157,4570,
                                          2719,4083,6586,4617,5137,7796,6564,
                                          7405,7298,7867,7580,9771,11647,
                                          25827,25616,15632,10454,13278,
                                          16858,27550,18719,48277,28639,32971,
                                          20762,17972,18975,15609,18617,
                                          6124,2116,3044,23544)
       )
```
# ggplot
```{r}
library(ggplot2)

baseplot <- ggplot(cdc) +
  aes(year, cases) +
  geom_point() +
  geom_line()

baseplot
```

> Q2. Using the ggplot geom_vline() function add lines to your previous plot for the 1946 introduction of the wP vaccine and the 1996 switch to aP vaccine (see example in the hint below). What do you notice?

```{r}
baseplot +
  geom_vline(xintercept = 1946, col="blue") +
  geom_vline(xintercept = 1995, col="red") +
  geom_vline(xintercept = 2020, col="gray") +
  geom_vline(xintercept = 2003, col="green")
```

> Q3. Describe what happened after the introduction of the aP vaccine? Do you have a possible explanation for the observed trend?

It is clear from the CDC data that pertussis cases are once again increasing. Vaccine effects wane so there is about a 10 year lag from the roll out in 1995.

# Exploring CMI-PB Data

```{r}
library(jsonlite)
subject <- read_json("https://www.cmi-pb.org/api/subject", simplifyVector = TRUE) 
head(subject, 3)

```

> Q4. How many aP and wP infancy vaccinated subjects are in the dataset?

```{r}
table(subject$infancy_vac)
```

> Q5. How many Male and Female subjects/patients are in the dataset? 

```{r}
table(subject$biological_sex)
```

> Q6. What is the breakdown of race and biological sex (e.g. number of Asian females, White males etc…)?

```{r}
table(subject$race, subject$biological_sex)
```
# Joining Multiple Tables
```{r}
specimen <- read_json("http://cmi-pb.org/api/v5/specimen",
                      simplifyVector = TRUE)
head(specimen)
```

> Q9. Complete the code to join specimen and subject tables to make a new merged data frame containing all specimen records along with their associated subject details:

```{r}
library(dplyr)
meta <- inner_join(subject, specimen)
head(meta)
```
# Expriment Data Table from CMI-PB

```{r}
abdata <- read_json("http://cmi-pb.org/api/v5/plasma_ab_titer",
                    simplifyVector = TRUE)
head(abdata)
```

> Q10. Now using the same procedure join meta with titer data so we can further analyze this data in terms of time of visit aP/wP, male/female etc.

```{r}
ab <- inner_join(abdata, meta)
head(ab)
```
> How many Ab measurements ?

```{r}
nrow(ab)
```

> Q11. How many specimens (i.e. entries in abdata) do we have for each isotype?

```{r}
table(ab$isotype)
```

> Antigens ?

```{r}
table(ab$antigen)
```

# Focusing on IgG
```{r}
igg <- filter(ab, isotype=="IgG")
head(igg)
```

> Q13. Complete the following code to make a summary boxplot of Ab titer levels (MFI) for all antigens:

```{r}
ggplot(igg) +
  aes(MFI_normalised, antigen) +
  geom_boxplot()
```
# adding color by aP/wP infancy_vac
```{r}
ggplot(igg) +
  aes(MFI_normalised, antigen, col=infancy_vac) +
  geom_boxplot()
```
# Another version of this plot adding infancy_vac to the faceting:
```{r}
ggplot(igg) +
  aes(MFI_normalised, antigen) +
  geom_boxplot() +
  facet_wrap(~visit)
```
```{r}
table(igg$visit)
```
Subjects with 8 visits or more lack data, so let's exclude
```{r}
igg_7 <- filter(igg, visit %in% 1:7)
table(igg_7$visit)
```

```{r}
ggplot(igg_7) +
  aes(MFI_normalised, antigen, col=infancy_vac) +
  geom_boxplot() +
  facet_wrap(~visit, ncol=2)
```
# Trying a different plot (x = time, y = MFI_Normalized)
```{r}
ggplot(igg_7) +
  aes(planned_day_relative_to_boost, MFI_normalised) +
  geom_point() +
  geom_line() 
```

```{r}
abdata.21 <- ab %>% filter(dataset == "2021_dataset")

abdata.21 %>% 
  filter(isotype == "IgG",  antigen == "PT") %>%
  ggplot() +
    aes(x=planned_day_relative_to_boost,
        y=MFI_normalised,
        col=infancy_vac,
        group=subject_id) +
    geom_point() +
    geom_line() 
```

