---
title: "Class16"
author: "Nicholas Thiphakhinkeo A17686679"
date: "2024-11-21"
output:
  pdf_document: default
  html_document: default
---

> Q. What does the star character accomplish here? Ask Barry, or your class neighbor, if you are not sure!

a powerful tool in shell commands to match multiple files or directories based on a pattern, in this case the .faa.gz

> Q. How many sequences are in this mouse.1.protein.faa file? Hint: Try using grep to figure this out…


> Q. What happens if you run the above command without the > mm-first.fa part?

If you run the command without the > mm-first.fa part, the first 11 lines of the mouse.1.protein.faa file will be printed directly to your terminal.

> Q. What happens if you were to use two ‘>’ symbols (i.e. >> mm-first.fa)?

Adds the new output to the end of the existing file.


# Reading Table
```{r}
b <- read.table("myresults.tsv", sep="\t", header=TRUE)
head(data)
```
# Setting Colnames
```{r}
colnames(b) <- c("qseqid", "sseqid", "pident", "length", "mismatch", "gapopen", "qstart", "qend", "sstart", "send", "evalue", "bitscore")
```

# Plot() Function
```{r}
plot(b$pident  * (b$qend - b$qstart), b$bitscore)
```
# GGPlot
```{r}
library(ggplot2)
ggplot(b, aes(pident, bitscore)) + geom_point(alpha=0.1) 
```
```{r}
ggplot(b, aes((b$pident * (b$qend - b$qstart)), bitscore)) + geom_point(alpha=0.1) + geom_smooth()
```


