---
title: "Class06"
author: Nicholas Thiphakhinkeo A17686679
format: pdf
---

#every fxn in R has atleast 3 things
# - a **name**, one or more input              **arguments**, The **body**, where 
```{r}
add <- function(x,y) {x+y}

```
```{r}
add(x=c(10,1,1,10),y=1)
```
```{r}
mean(c(10,10,NA), na.rm=T)
```
#Q1. Write a function grade() to determine an overall grade from a vector of student homework assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an NA value) this can be used as a score to be potentially dropped. Your final function should be adquately explained with code comments and be able to work on an example class gradebook such as this one in CSV format: “https://tinyurl.com/gradeinput” [3pts]
------------------------------
```{r}
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
```
# Student 1 Avg
```{r}
mean(student1)
```
# Student 2 Avg
```{r}
mean(student2, na.rm=T)
```
# Student 3
```{r}
mean(student3, na.rm=T)
```
# Subtracting Value by Eye
```{r}
mean(student1[-8])
```
# Min Function
```{r}
min.ind <- which.min(student1)
mean(student1[-min.ind])

```
```{r}
min.ind <- which.min(student2)
mean(student2[-min.ind], na.rm=F)
```
# Making NAs Zero
```{r}
student2[ is.na(student2) ] = 0
student2
```
# Finds min value bf getting mean
```{r}
grade <- function(x) {
  x[ is.na(x) ] <- 0
  mean(x[-which.min(x)])}
```
# Question 1: Writing Grade() Function
```{r}
gradebook <- read.csv("https://tinyurl.com/gradeinput",row.names=1)
head(gradebook)
ans <- apply(gradebook,1,grade)
```
# Question 2: Highest Scoring Student
```{r}
which.max(ans)
ans[which.max(ans)]
```
# Question 3: Which was toughest HW Assignment
```{r}
masked_gradebook <- gradebook
masked_gradebook[ is.na(masked_gradebook)] = 0
apply(masked_gradebook,2,mean)
```
# Grade2
```{r}

```
# Q4: Cor()
```{r}
x <- c(100,90,80,100) 
y <- c(100,90,81,99)
z <- c(100,80,90,100)
cor(x,z)
```
```{r}
cor(ans,masked_gradebook$hw5)
```
```{r}
apply(masked_gradebook, 2, cor, y=ans)
```

