---
title: "Class07"
author: Nicholas Thiphakhinkeo
format: pdf
---
# rnorm
```{r}
rnorm(10)
```
```{r}
hist(rnorm(1000, mean=3))
```

```{r}
n <- 1000
x <- c(rnorm(n,-3),rnorm(n,+3))
hist(x)
```


```{r}
n<-30
x<-c(rnorm(n,-3), rnorm(n,+3))
y <- rev(x)
z <- cbind(x,y)
head(z)
```

```{r}
plot(z)
```

```{r}
km <- kmeans(z,centers=2)
km
```

```{r}
km$size
```
Cluster Assignment
```{r}
km$cluster
```
Cluster Center
```{r}
km$centers
```
# Plot z color colored by kmeans cluster assignment and add cluster centers as blue points
```{r}

```
R recycles shorter color vector to be the same length as the longer (number of data points) in z
```{r}
plot(z, col=c("purple","green"))
```

```{r}
plot(z, col=km$cluster)
```
# use points() function to add new points to an existing plot... like a cluster plot
```{r}
plot(z, col=km$cluster)
points(km$centers, col="purple", pch=15, cex=2)
```
# Q. Run kmeans and ask for 4 clusters and plot results
```{r}
km2 <- kmeans(z, centers=4)
km2
```

```{r}
plot(z, col=km2$cluster)
points(km2$centers, col="purple", pch=15, cex=1.5)
```
# Hierarchical Clustering 
#need distance matrix of data to be clustered
```{r}
d <- dist(z)
hc<- hclust(d)
hc
```

```{r}
plot(hc)
```
#Cluster Membership by cutting tree with cutree()
```{r}
grps <- cutree(hc,h=8)
grps
```
# Plot "z" colored by hclust 
```{r}
plot(z, col=grps)
```
# PCA Uk Foods Data
```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url, row.names=1)
head(x)
```

```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```

```{r}
pairs(x, col=rainbow(10), pch=16)
```
# Using PCA for Larger Data Sets
```{r}
pca <- prcomp(t(x))
summary(pca)
```

```{r}
attributes(pca)
```

```{r}
pca$x
```

```{r}
plot(pca$x[,1],pca$x[,2], col=c("black","red","blue","darkgreen"), pch=15, xlab="PC1 (67.4%)", ylab="PC2(29.05%)")
```
# Colored Country Plot
```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1(67.4%)", ylab="PC2(29.05%)", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x),col=c("yellow","red","blue","darkgreen"),pch=0, cex=1)
```

