---
title: "Class10"
author: Nicholas Thiphakhinkeo A17686679
format: pdf
---

# Read/Inspect
```{r}
candy_file <- "candy-data.csv"
candy = read.csv(candy_file, row.names=1)
head(candy)
```
> Q1. How many different candy types are in this dataset?

```{r}
num_candy_types <- nrow(candy)
print(num_candy_types)
```

> Q2. How many fruity candy types are in the dataset?

```{r}
fruity_candy <- candy[candy$fruity==T,]
num_fruity_candy_types <- nrow(fruity_candy)
print(num_fruity_candy_types)
```

> Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
candy["Sour Patch Kids", ]$winpercent
```

> Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat", ]$winpercent
```

> Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```
# Skim Function
```{r}
library("skimr")
skim(candy)
```

> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

sugarpercent, pricepercent, and winpercent show percent rather than value.

> Q7. What do you think a zero and one represent for the candy$chocolate column?

The 0s are likely for candies not containing chocolate while is are for candies that do.

> Q8. Plot a histogram of winpercent values

```{r}
library("ggplot2")
ggplot(candy, aes(x = winpercent)) +
  geom_histogram(fill = "purple", color = "black") +
    geom_vline(xintercept = 50, color = "red") +
  labs(title = "Distribution of Win Percentage", x = "Win Percentage", y = "Frequency")
```
> Q9. Is the distribution of winpercent values symmetrical?

No the distribution is right-skewed

> Q10. Is the center of the distribution above or below 50%?

Below 50% suggesting most candies have win percentage below 50%

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
mean_chocolate_winpercent <- mean(candy$winpercent[as.logical(candy$chocolate)])
mean_fruity_winpercent <- mean(candy$winpercent[as.logical(candy$fruity)])

print(mean_chocolate_winpercent)
print(mean_fruity_winpercent)
```
Chocolate is ranked higher than fruity

> Q12. Is this difference statistically significant?

```{r}
t_test1 <- t.test(candy$winpercent[as.logical(candy$chocolate)], candy$winpercent[as.logical(candy$fruity)])
print(t_test1)
```
p value <-0.05 so likely significant

> Q13.  What are the five least liked candy types in this set?

```{r}
head(candy[order(candy$winpercent),], n = 5)
```
> Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
head(candy[order(candy$winpercent, decreasing = TRUE),], n = 5)
```

> Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
ggplot(candy) + 
  aes(x = rownames(candy), y = winpercent) + 
  geom_col() +
  coord_flip() + 
  labs(title = "Candy Ranking by Win Percentage",
       x = "Candy Name",
       y = "Win Percentage")
```

> Q16. use the reorder() function to get the bars sorted by winpercent?

```{r}
ggplot(candy) + 
  aes(x = reorder(rownames(candy), winpercent), y=winpercent) + 
  geom_col() +
  coord_flip() + 
  labs(title = "Candy Ranking by Win Percentage",
       x = "Candy Name",
       y = "Win Percentage")
```
# Adding Color
```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"

ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```

> Q17. What is the worst ranked chocolate candy?

The worst ranked chocolate candies are Sixlets

> Q18. What is the best ranked fruity candy?

The best ranked fruity candy is Starburst

# Taking a look at pricepercent 
```{r}
library("ggrepel")
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 78)
```

> Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

Candy Bars

> Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
library("dplyr")
candy %>% 
 arrange(desc(pricepercent)) %>% 
 head(5)
```
Nik L Nips are the most expensive and least popular 

# Corrplot
```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```
> Q22. Examining this plot what two variables are anti-correlated (i.e.have minus values)?

Fruity winpercentage is least correlated

> Q23. Similarly, what two variables are most positively correlated?

Chocolate winpercentage is most positively correlated 

# PCA

```{r}
pca <- prcomp(candy, scale = TRUE)
summary(pca)
```
# PC1vPC2
```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```
# PCAggplot
```{r}
my_data <- cbind(candy, pca$x[,1:3])
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)
p
```
# PCAggrepel
```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 53)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

# PCAplotly
```{r}
library(plotly)
```
# PCAloadings
```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

> Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

Fruity, Hard, and Pluribus are picked up strongly in the positive direcetion. Likely since they have lower win percentages it is easier to sell them in bulk compared to singularly.